Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-17 VMCAI-revised-version/final.tex
up

\begin{proof}
\alain{preuve pas claire, à réécrire}
\alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
\mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input in $\uparrow(0,0,1)$ 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
Remark that in the construction, from any input in $\uparrow(0,0,1)$, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
Hence {\sc Resilience} is undecidable for Minsky machines.
\alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
\alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS such that
 an execution from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS simulates
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 a state of the form $(s',0,0,0)$, and
 correspondingly,
  if there is a sequence from 
  $(s_0,0,0,0)$ to 
  $(s_f,0,0,0)$ in the Minsky machine
  there is a sequence in the reset-VASS which simulates
  the sequence and ends
  in a state of the form $(s,0, \ldots O)$
  (from Theorem~$5$ in \cite{araki1976PN}).
Hence $\downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
resilience in the corresponding reset-VASS.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, $\post^*$-basis $\omega^2$-WSTS with strong compatibility as well.\qed
%\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
\end{proof}

>>>>

\begin{proof}
\alain{preuve pas claire, à réécrire}
\alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
\mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input in $\uparrow(0,0,1)$ 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
Remark that in the construction, from any input in $\uparrow(0,0,1)$, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
Hence {\sc Resilience} is undecidable for Minsky machines.
\alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
\alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS such that
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 a state of the form $(s',0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
Hence $\downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
resilience in the corresponding reset-VASS.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, $\post^*$-basis $\omega^2$-WSTS with strong compatibility as well.\qed
%\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
\end{proof}
